---
title: "Week 3 Assignment"
author: "Joshua Trahan"
date: "11/12/2021"
output:
  pdf_document: default
  html_document: default
---

```{r import, include=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(chron)
library(lubridate)
url <- 'https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD'
nypd_df <- read.csv(url)
```

# New York City: Analysis of Shooting Incidents
> The dataset includes all of the shooting incidents that occured in NY City from 2006 until the end of 2020.

```{r clean, include=TRUE, message=FALSE, warning=FALSE}
nypd_df <- nypd_df %>%
  select(-c(INCIDENT_KEY, JURISDICTION_CODE, LOCATION_DESC)) %>%
  mutate(OCCUR_DATE = mdy(OCCUR_DATE), PRECINCT = as.factor(PRECINCT),
         STATISTICAL_MURDER_FLAG = as.logical(STATISTICAL_MURDER_FLAG),
         OCCUR_TIME = chron(times = OCCUR_TIME))
```

## The chart below depicts the number of murders by New York City boro
>Statin Island and Queen have reported the fewest while Brooklyn and the Bronx the most.

```{r boro_murders, include=TRUE, message=FALSE, warning=FALSE}
boro_shootings_deaths <- nypd_df %>%
  select(OCCUR_DATE, BORO, STATISTICAL_MURDER_FLAG)%>%
  mutate(STATISTICAL_MURDER_FLAG = as.integer(nypd_df$STATISTICAL_MURDER_FLAG))

boro_shootings_deaths <- boro_shootings_deaths %>%
  mutate(MM_YY = format(as.Date(OCCUR_DATE), '%m-%y'))
boro_shootings_deaths1 <- boro_shootings_deaths %>%
  mutate(Year = format(as.Date(OCCUR_DATE),'%y'))
  

```
```{r, stack_bar, include=TRUE, message=FALSE, warning=FALSE}
ggplot(boro_shootings_deaths1, aes(fill=BORO, y=STATISTICAL_MURDER_FLAG, x = Year))+
  geom_bar(position="stack", stat="identity")+
  labs(title = 'Boro Deaths by Month', x = 'Year', y = 'Deaths')


view(boro_shootings_deaths)
```

## Data Summary
*Observations are missing in perpetrator data*

>This data will be omitted for logistic regression.
>Rows are included for accurate shooting and death comparisons.


```{r shoots_v_deaths, include=TRUE, message=FALSE, warning=FALSE}
NY_shooting_v_Death <- boro_shootings_deaths %>%
  mutate(Shooting = is.integer(boro_shootings_deaths$STATISTICAL_MURDER_FLAG))
NY_shooting_v_Death <- NY_shooting_v_Death %>%
  select(MM_YY, STATISTICAL_MURDER_FLAG, Shooting) %>%
  mutate(MM_YY = factor(NY_shooting_v_Death$MM_YY))%>%
  group_by(MM_YY) %>%
  summarize(STATISTICAL_MURDER_FLAG = sum(STATISTICAL_MURDER_FLAG), Shooting = sum(as.numeric(Shooting)))
str(NY_shooting_v_Death)
summary(NY_shooting_v_Death)
  
colnames(NY_shooting_v_Death)[2] <- 'Murder'
colnames(NY_shooting_v_Death)[1] <- 'Date'
  
```
```{r shoot_v_death_line, include=TRUE, message=FALSE, warning=FALSE}
NY_shooting_v_Death_plot <- NY_shooting_v_Death %>%
  ggplot(aes(x=Date, y=Shooting, group=1))+
  geom_line(aes(color = 'Shooting'))+
  geom_line(aes(y=Murder, color = 'Murder'))+geom_point(aes(color='Shooting'))+
  geom_point(aes(y=Murder, color = 'Murder'))+
  scale_y_log10()
```

## Trends between shooting incidents and corresponding death


```{r Graph, include=TRUE, message=FALSE, warning=FALSE}
NY_shooting_v_Death_plot
```



## Simple Linear Regression
>This regression analysis depicts the relationship between shootings and shooting deaths. Shooting deaths increase along with shootings. As the number of shootings increase it becomes more difficult to predict resulting deaths. Analysis of outliers may help with understanding the nature of shootings.

```{r shoot_death_lm, include=TRUE, message=FALSE, warning=FALSE}
lm_death <- lm(Murder ~ Shooting, data = NY_shooting_v_Death)
plot(x = NY_shooting_v_Death$Shooting, y= NY_shooting_v_Death$Murder, xlab = 'Shootings', ylab = 'Deaths')+
  abline(lm_death)+
  geom_text(aes(y = Deaths, label = Murder), size = 4)
```
## Identifying Bias

### In order to avoid a bias dataset, additional data is needed about the population of the provided areas. The data cannot answer the *why*. No additional information is provided about the precincts, state economic investment, or detailed demographics. 
